Scikit-learn: API design for machine learning

### 2013-06-20 paper.tex
wordings (2., 2.1, 2.2)

\begin{description}
  \item[Consistency.]
       All objects (basic or composed) share a consistent interface composed of
       a limited set of methods. This interface is documented in a consistent
       manner for all objects.
  \item[Inspection.]
       Constructor parameters and parameter values determined by learning
       algorithms are stored and exposed as public attributes.
  \item[Non-proliferation of classes.]
       Learning algorithms are the only objects to be of custom classes.
       Datasets and hyper-parameter names and values are represented as standard
       Python, NumPy or SciPy classes rather than new ones whenever feasible.
       This keeps \sklearn easy to use and easy to combine with other libraries.
       % XXX do we need to elaborate more on this, or do we expect everyone
       % to be an experienced Java/adapter pattern hater?
  \item[Composition.]
       Many machine learning tasks are expressible
       as sequences or combinations of transformations to data.
       Some learning algorithms are also naturally viewed
       as meta-algorithms parametrized on other algorithms.
       Whenever possible, those algorithms are implemented and composed from
       existing building blocks.
  \item[Sensible defaults.]
       Whenever an operation requires a user-defined parameter,
       a default and appropriate value is defined by the library.
       The default value should cause the operation to be performed
       in a sensible way (giving a baseline solution for the task at hand).
\end{description}

@@

In \sklearn, we chose a representation of data that is as close as
possible to the matrix representation: datasets are encoded as NumPy
multidimensional arrays for dense data and as SciPy sparse matrices for sparse
data. While those may seem rather unsophisticated data representations when
compared to more object-oriented constructs, such as those used by
Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
% Lars: "orders of magnitude faster" is an implementation detail
on all NumPy and SciPy vector operations while allowing us at the same time to
keep the code short and readable.  This design choice has also been motivated by
the fact that, given their pervasiveness in many other scientific Python
packages, many scientific users of Python are already familiar with NumPy dense
arrays and SciPy sparse matrices.
% Gilles: the phrase was too long
% hence helping them to familiarize with the project.
From a practical point of view, these formats also provide a bunch of
data loading and conversion tools which make them very easy to use in many
contexts. Moreover, for tasks where the inputs are text files or semi-structured
objects, we provide \textit{vectorizer} objects that efficiently convert such
data to the NumPy or SciPy formats.

@@

Actual learning is performed by the \texttt{fit} method. This method is called
with training data (e.g., supplied as two arrays \texttt{X\_train} and
\texttt{y\_train} in supervised learning estimators). Its task is to run a
learning algorithm and to determine model-specific parameters from the training
data and set these as attributes on the estimator object. As a convention, the
parameters learned by an estimator are exposed as public attributes whose name
is suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
learned coefficients of a linear model),
again to facilitate model inspection.
In the partial application view,
\texttt{fit} is a function from data to a model of that data.
As such, it always returns the estimator object it was called on,
which now serves as a model of its input and can be used to perform predictions or transformations of input data.
The choice to let a single object serve dual purpose as estimator and model
simplifies the \sklearn class hierarchy
compared to an approach were estimators and models are separated.

To illustrate the initialize-\texttt{fit} sequence,
let us consider a supervised learning task using logistic regression.
Given the API defined above, solving this problem is as simple as:\footnote{
  \texttt{import} statement are omitted for brevity
  throughout this paper.}
\begin{pythoncode}
clf = LogisticRegression(penalty="l1")
clf.fit(X_train, y_train)
\end{pythoncode}
In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
setting the \texttt{penalty} hyper-parameter to \texttt{"l1"}, hence enabling
$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C}, which
defines the strength of the regularization) are not explicitly defined and are
thus set to the default values. Upon calling \texttt{fit}, the model is then
learned from the training arrays \texttt{X\_train} and \texttt{y\_train}.
(In this example, \texttt{fit} is called solely for its side effects.)
Since all estimators share the same interface, using a different learning algorithm is
as simple as replacing the constructor (the class name).
For example, to build a random forest on
the same training data, one would simply have to replace
\texttt{LogisticRegression(penalty="l1")} in the snippet above by \\
\texttt{RandomForestClassifier()}.

>>>>

\begin{description}
  \item[Consistency.]
       All objects (basic or composite) share a consistent interface composed of
       a limited set of methods. This interface is documented in a consistent
       manner for all objects.
  \item[Inspection.]
       Constructor parameters and parameter values determined by learning
       algorithms are stored and exposed as public attributes.
  \item[Non-proliferation of classes.]
       Learning algorithms are the only objects to be of custom classes.
       Datasets and hyper-parameter names and values are represented as standard
       Python, NumPy or SciPy classes rather than new ones whenever feasible.
       This keeps \sklearn easy to use and easy to combine with other libraries.
  \item[Composition.]
       Many machine learning tasks are expressible
       as sequences or combinations of transformations to data.
       Some learning algorithms are also naturally viewed
       as meta-algorithms parametrized on other algorithms.
       Whenever possible, those algorithms are implemented and composed from
       existing building blocks.
  \item[Sensible defaults.]
       Whenever an operation requires a user-defined parameter,
       a default and appropriate value is defined by the library.
       The default value should cause the operation to be performed
       in a sensible way (giving a baseline solution for the task at hand).
\end{description}

@@

In \sklearn, we chose a representation of data that is as close as
possible to the matrix representation: datasets are encoded as NumPy
multidimensional arrays for dense data and as SciPy sparse matrices for sparse
data. While those may seem rather unsophisticated data representations when
compared to more object-oriented constructs, such as those used by
Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
on all NumPy and SciPy vector operations while keeping
the code short and readable.  This design choice has also been motivated by
the fact that, given their pervasiveness in many other scientific Python
packages, many scientific users of Python are already familiar with NumPy dense
arrays and SciPy sparse matrices.
From a practical point of view, these formats also provide a bunch of
data loading and conversion tools which make them very easy to use in many
contexts. Moreover, for tasks where the inputs are text files or semi-structured
objects, we provide \textit{vectorizer} objects that efficiently convert such
data to the NumPy or SciPy formats.

@@

Actual learning is performed by the \texttt{fit} method. This method is called
with training data (e.g., supplied as two arrays \texttt{X\_train} and
\texttt{y\_train} in supervised learning estimators). Its task is to run a
learning algorithm and to determine model-specific parameters from the training
data and set these as attributes on the estimator object. As a convention, the
parameters learned by an estimator are exposed as public attributes whose names
are suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
learned coefficients of a linear model),
again to facilitate model inspection.
In the partial application view,
\texttt{fit} is a function from data to a model of that data.
As such, it always returns the estimator object it was called on,
which now serves as a model of its input and can be used to perform predictions or transformations of input data.
The choice to let a single object serve dual purpose as estimator and model
simplifies the \sklearn class hierarchy
compared to an approach were estimators and models are separated.

To illustrate the initialize-\texttt{fit} sequence,
let us consider a supervised learning task using logistic regression.
Given the API defined above, solving this problem is as simple as:\footnote{
  \texttt{import} statements are omitted for brevity
  throughout this paper.}
\begin{pythoncode}
clf = LogisticRegression(penalty="l1")
clf.fit(X_train, y_train)
\end{pythoncode}
In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
setting the \texttt{penalty} hyper-parameter to \texttt{"l1"}, hence enabling
$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C}, which
defines the strength of the regularization) are not explicitly defined and are
thus set to the default values. Upon calling \texttt{fit}, the model is then
learned from the training arrays \texttt{X\_train} and \texttt{y\_train}.
(In this example, \texttt{fit} is called solely for its side effects.)
Since all estimators share the same interface, using a different learning algorithm is
as simple as replacing the constructor (the class name).
For example, to build a random forest on
the same training data, one would simply have to replace
\texttt{LogisticRegression(penalty="l1")} in the snippet above by \\
\texttt{RandomForestClassifier()}.
